### 2019-04-16 papers/haskell19/2-sources.tex
Code refactor

\section{Sources of Constructions}
\label{sec:sources}

@@

\begin{code}
data Html
  =  Text  String
  |  Sing  String
  |  Tag   String Html
  |  Html  :+: Html
\end{code}

@@

(<+>) :: Html -> Html -> Html
(<+>) x y = x :+: br :+: y

list :: [Html] -> Html
list xs = Tag "ul" (foldl1 (:+:) (Tag "li" <$> xs))
\end{code} %$

\begin{code}
simplify :: Html -> Html
simplify (Text t1 :+: Text t2)
  = Text (t1 ++ t2)
simplify ((Text t1 :+: x) :+: y)
  = simplify (Text t1 :+: simplify (x :+: y))
simplify (x :+: y)
  = simplify x :+: simplify y
simplify (Tag t x)
  = Tag t (simplify x)
simplify x = x
\end{code}

>>>>

\section{Random Testing using Hints}
\label{sec:sources}

@@

\begin{code}
data Html
  =  Text  String
  |  Sing  String
  |  Tag   String Html
  |  Join  Html Html
\end{code}

A value of type |Html| represents either plain text strings, via the |Text| data
constructor; single and paired HTML tags using the |Sing| and |Tag| data
constructors, respectively; and concatenation of ...

@@

(<+>) :: Html -> Html -> Html
(<+>) x y = Join x (Join br y)

list :: [Html] -> Html
list xs = Tag "ul" (foldl1 Join (Tag "li" <$> xs))
\end{code} %$

\begin{code}
simplify :: Html -> Html
simplify (Join (Text t1) (Text t2))
  = Text (t1 ++ t2)
simplify (Join (Join (Text t1) x) y)
  = simplify (Join (Text t1) (simplify (Join x y)))
simplify (Join x y)
  = Join (simplify x) (simplify y)
simplify (Tag t x)
  = Tag t (simplify x)
simplify x = x
\end{code}

### 2019-04-16 papers/haskell19/3-representation.tex
Code refactor

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
Nam vestibulum accumsan nisl.

@@

% cata :: Functor f => (f a -> a) -> Fix f -> a
% cata f = f . fmap (cata f) . unFix

\begin{code}
data Con_Text   r = Con_Text String
data Con_Sing   r = Con_Sing String
data Con_Tag    r = Con_Tag  String r
data Con_infix  r = Con_infix r r
\end{code}

\begin{code}
data Fun_br     r = Fun_br
data Fun_bold   r = Fun_bold r
data Fun_infix  r = Fun_infix r r
data Fun_list   r = Fun_list [r]
\end{code}

@@

instance Algebra Con_infix Html where
  alg (Con_infix x y) = x :+: y
\end{code}

@@

instance Algebra Fun_infix Html where
  alg (Fun_infix x y) = x <+> y

@@

\begin{code}
instance Algebra Pat_simplify_1 Html where
  alg (Pat_simplify_1 t1 t2) = Text t1 :+: Text t2

instance Algebra Pat_simplify_2 Html where
  alg (Pat_simplify_1 t x y) = (Text t :+: x) :+: y
\end{code}

@@

\begin{code}
type family Rep (t :: Symbol) :: * -> *

type family Con (c :: Symbol) :: * -> *
type family Fun (f :: Symbol) :: * -> *
type family Pat (p :: Symbol) (n :: Nat) :: * -> *
\end{code}

>>>>

% --------------------------------------
% The basic idea
%
This section introduces a unified representation for the different constructions
we might want to consider when generating random values.
%
The key idea of this work is to lift each different source of structural
information to the type level.
%
In this light, we will define a set of different simple ``open'' data types,
each one in charge of representing a single construction.
%
Later, these types can be
\begin{inparaenum}[(i)]
  \item combined in several ways depending on the desired shape of our test data,
\item randomly generated (see Section \ref{sec:generators}), and finally,
\item transformed back to corresponding values of the target data type automatically.
\end{inparaenum}
%
This lifted representation of constructions can be automatically derived from
our source code in compile-time, relieving programmers of the burden of writing
random generators by hand.

% --------------------------------------
% Representing data constructors
%
\paragraph{Representing Data Constructors}
%
The simplest piece of meaningful%
\footnote{We avoid considering $\bot$ as it does not contain any useful
  information.}
%
information we can consider when generating values of an algebraic data type, is
the one given by each one of its data constructors.
%
In this light, the representation of each individual data constructor of the
target data type is simply given by a new data type containing a single
constructor with the same fields as the corresponding one, except for the
recursive fields which are abstracted away using a type parameter.

Recalling our |Html| example, each one of it data constructors can be
represented as follows:

\begin{code}
data Con_Text   r = Con_Text String
data Con_Sing   r = Con_Sing String
data Con_Tag    r = Con_Tag  String r
data Con_Join   r = Con_Join r r
\end{code}

@@

@@

instance Algebra Con_Join Html where
  alg (Con_Join x y) = Join x y
\end{code}

\begin{code}
data Fun_br     r = Fun_br
data Fun_bold   r = Fun_bold r
data Fun_cat    r = Fun_cat r r
data Fun_list   r = Fun_list [r]
\end{code}

@@

instance Algebra Fun_cat Html where
  alg (Fun_cat x y) = x <+> y

@@

\begin{code}
instance Algebra Pat_simplify_1 Html where
  alg (Pat_simplify_1 t1 t2) = Join (Text t1) (Text t2)

instance Algebra Pat_simplify_2 Html where
  alg (Pat_simplify_1 t x y) = Join (Join (Text t) x) y
\end{code}

@@

\begin{code}
type family Con (c :: Symbol)             :: * -> *
type family Fun (f :: Symbol)             :: * -> *
type family Pat (p :: Symbol) (n :: Nat)  :: * -> *
\end{code}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec hendrerit tempor
tellus. Donec pretium posuere tellus. Proin quam nisl, tincidunt et, mattis
eget, convallis nec, purus. Cum sociis natoque penatibus et magnis dis
parturient montes, nascetur ridiculus mus. Nulla posuere. Donec vitae dolor.
Nullam tristique diam non turpis. Cras placerat accumsan nulla. Nullam rutrum.
Nam vestibulum accumsan nisl\cite{godefroid2005dart}.

### 2019-04-16 papers/haskell19/4-generators.tex
Code refactor

\section{Generating Composable Data via Type Specfications}
\label{sec:generators}

\begin{code}
type BoundedGen a = Int -> Gen a

class BoundedArb (a :: *) where
  boundedArb :: BoundedGen a
\end{code}

@@

\begin{code}
class BoundedArb1 (f :: * -> *) where
  liftBounded :: BoundedGen a -> BoundedGen (f a)
\end{code}

@@

\begin{code}
instance BoundedArb1 f => BoundedArb1 (f otimes n) where
  liftBounded gen lvl = Freq <$> liftBounded gen lvl

instance BoundedArb1 f => BoundedArb1 (Term f) where
  liftBounded gen lvl = Term <$> liftBounded gen lvl
\end{code}

@@

\begin{code}
instance (BoundedArb1 f, BoundedArb1 g)
  => BoundedArb1 (f oplus g) where
  liftBounded gen lvl =
    if lvl > 0
    then frequency
    [ (freqVal  at(FreqOf f) ,  InL <$> liftBounded gen lvl)
    , (freqVal  at(FreqOf g) ,  InR <$> liftBounded gen lvl) ]
    else frequency
    [ (freqVal  at(FreqOf' f) ,  InL <$> liftBounded gen lvl)
    , (freqVal  at(FreqOf' g) ,  InR <$> liftBounded gen lvl) ]
\end{code}

\begin{code}
instance BoundedArb1 Con_Text where
  liftBounded gen lvl = Con_Text <$> arbitrary

instance BoundedArb1 Con_Sing where
  liftBounded gen lvl = Con_Sing <$> arbitrary

instance BoundedArb1 Con_Tag where
  liftBounded gen lvl
    = Con_Tag <$> arbitrary <*> gen (lvl - 1)

instance BoundedArb1 Con_infix where
  liftBounded gen lvl
    = Con_infix <$> gen (lvl - 1) <*> gen (lvl - 1)
\end{code} %$

\begin{code}
instance BoundedArb1 Fun_br where
  liftBounded gen lvl = pure Fun_br

instance BoundedArb1 Fun_bold where
  liftBounded gen lvl = Fun_bold <$> gen (lvl - 1)

instance BoundedArb1 Fun_infix where
  liftBounded gen lvl
    = Fun_infix <$> gen (lvl - 1) <*> gen (lvl - 1)

instance BoundedArb1 Fun_list where
  liftBounded gen lvl
    = Fun_list <$> liftGen (gen (lvl - 1))
\end{code} %$

\begin{code}
instance BoundedArb1 Pat_simplify_1 where
  liftBounded gen lvl
    = Pat_simplify_1 <$> arbitrary <*> arbitrary

instance BoundedArb1 Pat_simplify_2 where
  liftBounded gen lvl
    = Pat_simplify_2 <$> gen (lvl - 1) <*> gen (lvl - 1)
\end{code}

\begin{code}
instance BoundedArb1 f => BoundedArb (Fix f) where
  boundedArb lvl = Fix <$> liftBounded boundedArb lvl
\end{code} %$

\begin{code}
genFix  ::  forall f dot BoundedArb1 f =>  BoundedGen (Fix f)
genFix lvl = boundedArb at(Fix f) lvl
\end{code}

\begin{code}
genEval :: forall f a dot (BoundedArb1 f, Algebra f a) =>  BoundedGen a
genEval lvl = eval <$> genFix at f lvl
\end{code} %$

>>>>

\section{Generating Composable Data via Type-level Specfications}
\label{sec:generators}

\begin{code}
type BGen a = Int -> Gen a

class BArbitrary (a :: *) where
  bounded :: BGen a
\end{code}

@@

\begin{code}
class BArbitrary1 (f :: * -> *) where
  liftBGen :: BGen a -> BGen (f a)
\end{code}

@@

\begin{code}
instance BArbitrary1 f => BArbitrary1 (f otimes n) where
  liftBGen gen lvl = Freq <$> liftBGen gen lvl

instance BArbitrary1 f => BArbitrary1 (Term f) where
  liftBGen gen lvl = Term <$> liftBGen gen lvl
\end{code}

@@

\begin{code}
instance (BArbitrary1 f, BArbitrary1 g)
  => BArbitrary1 (f oplus g) where
  liftBGen gen lvl =
    if lvl > 0
    then frequency
      [ (freqVal  at(FreqOf f) ,  InL <$> liftBGen gen lvl)
      , (freqVal  at(FreqOf g) ,  InR <$> liftBGen gen lvl) ]
    else frequency
      [ (freqVal  at(FreqOf' f) ,  InL <$> liftBGen gen lvl)
      , (freqVal  at(FreqOf' g) ,  InR <$> liftBGen gen lvl) ]
\end{code}

\begin{code}
instance BArbitrary1 Con_Text where
  liftBGen gen lvl = Con_Text <$> arbitrary

instance BArbitrary1 Con_Sing where
  liftBGen gen lvl = Con_Sing <$> arbitrary

instance BArbitrary1 Con_Tag where
  liftBGen gen lvl = Con_Tag <$> arbitrary <*> gen (lvl - 1)

instance BArbitrary1 Con_Join where
  liftBGen gen lvl = Con_Join <$> gen (lvl - 1) <*> gen (lvl - 1)
\end{code} %$

\begin{code}
instance BArbitrary1 Fun_br where
  liftBGen gen lvl = pure Fun_br

instance BArbitrary1 Fun_bold where
  liftBGen gen lvl = Fun_bold <$> gen (lvl - 1)

instance BArbitrary1 Fun_cat where
  liftBGen gen lvl = Fun_cat <$> gen (lvl - 1) <*> gen (lvl - 1)

instance BArbitrary1 Fun_list where
  liftBGen gen lvl = Fun_list <$> liftGen (gen (lvl - 1))
\end{code} %$

\begin{code}
instance BArbitrary1 Pat_simplify_1 where
  liftBGen gen lvl
    = Pat_simplify_1 <$> arbitrary <*> arbitrary

instance BArbitrary1 Pat_simplify_2 where
  liftBGen gen lvl
    = Pat_simplify_2 <$> gen (lvl - 1) <*> gen (lvl - 1)
\end{code}

\begin{code}
instance BArbitrary1 f => BArbitrary (Fix f) where
  bounded lvl = Fix <$> liftBGen bounded lvl
\end{code} %$

\begin{code}
genFix  ::  forall f dot BArbitrary1 f =>  BGen (Fix f)
genFix lvl = bounded at(Fix f) lvl
\end{code}

\begin{code}
genEval :: forall f a dot (BArbitrary1 f, Algebra f a) => BGen a
genEval lvl = eval <$> genFix at f lvl
\end{code} %$
